Cover Times of Random Walks

### 2019-04-20 paper.tex
spanning tree argument and binary trees

\DeclareMathOperator{\E}{\mathrm{E}}		     % expected value
\DeclareMathOperator{\pr}{\mathrm{P}}		     % expected value
\DeclareMathOperator{\cov}{\uptau_\textrm{cov}}  % cover time variable
\DeclareMathOperator{\tcov}{t_\textrm{cov}}      % cover time
\DeclareMathOperator{\hit}{t_{\textrm{hit}}}     % hitting time
\DeclareMathOperator{\tmina}{t_\textrm{min}^A}   % minimum time on A
\DeclareMathOperator{\Ceff}{C_\textrm{eff}}      % effective conductance
\DeclareMathOperator{\Reff}{R_{\textrm{eff}}}    % effective resistance

\newcommand{\todo}[1]{{\color{blue}{[{\bf TODO:} #1]}}}

@@

\subsection{Definitions}
We consider an undirected graph $G$ without self loops or multi-edges.
Let $V(G)$ be the set of vertices and let $E(G)$ be the set of edges on $G$.
Then $n$ is the number of vertices $|V(G)|$.
%Without loss of generality, we label the vertices $\{1, 2, 3, ..., n\}$.

We say $i\sim j$ for $i$ and $j$ in the vertex set $V(G)$ if $(i,j)$ in $E(G)$.
Let $C_{i,j}$ be the non-negative weight of the edge between vertices $i$ and $j$.
We will mostly consider unweighted graphs
with $C_{i,j} = 1$ if $i \sim j$ and $C_{i,j} = 0$ otherwise.

@@

\subsection{Commute Time Identity}
We use the electrical network analysis in \cref{sec:electric} to 
prove the commute time identity so we may use it to help us bound
cover times.

@@

\subsection{Matthews Method on Balanced Trees}
\cite{LP17} 11.3

\subsection{Spanning Tree Argument}
\cite{AF14} 3.2

Edge commute inequality

spanning trees

@@

>>>>

\DeclareMathOperator{\E}{\mathrm{E}}		     % expected value
\DeclareMathOperator{\pr}{\mathrm{P}}		     % expected value
\DeclareMathOperator{\cov}{\uptau_\textrm{cov}}  % cover time variable
\DeclareMathOperator{\tcov}{t_\textrm{cov}}      % cover time
\DeclareMathOperator{\hit}{t_{\textrm{hit}}}     % hitting time
\DeclareMathOperator{\tmina}{t_\textrm{min}^A}   % minimum time on A
\DeclareMathOperator{\Ceff}{C_\textrm{eff}}      % effective conductance
\DeclareMathOperator{\Reff}{R_{\textrm{eff}}}    % effective resistance
\DeclareMathOperator{\Rxy}{R_{\textrm{x,y}}}     % effective resistance between x,y
\DeclareMathOperator{\T}{\mathcal{T}}            % spanning tree

\newcommand{\todo}[1]{{\color{blue}{[{\bf todo:} #1]}}}

@@

\subsection{Definitions}
We consider an undirected, connected graph $G$ without self loops or multi-edges.
Let $V(G)$ be the set of vertices and let $E(G)$ be the set of edges on $G$.
Then $n$ is the number of vertices $|V(G)|$.
We say $i\sim j$ for $i$ and $j$ in the vertex set $V(G)$ if $(i,j)$ in $E(G)$.
Assume that each vertex is accessible from every other vertex.
That is, there is a set of edges from any $x$ to any $y$.
%Without loss of generality, we label the vertices $\{1, 2, 3, ..., n\}$.

Let $C_{i,j}$ be the non-negative weight of the edge between vertices $i$ and $j$.
We will mostly consider unweighted graphs
with $C_{i,j} = 1$ if $i \sim j$ and $C_{i,j} = 0$ otherwise.

@@

\subsection{Commute Time Identity} \label{sec:commute}
We use the electrical network analysis in \cref{sec:electric} to 
prove the commute time identity so we may use it to help us bound
cover times.

@@

\subsection{Matthews Method on Binary Trees}
\todo{graph example of binary tree}

\cite{LP17}
Consider a complete unweighted binary tree with a single root $\rho$.
Each vertex has two children if it is distance less than $k$
from the root and no children if it is a leaf (at distance $k$ from the root).
There are $n = 1 + 2 + 4 + ... + 2^k = 2^{k+1} - 1$ vertices.

By \cref{eqn:commute}, the commute time between the root  and a leaf $a$ is
\begin{align}
C R_{\rho, a} = 2 (n-1) k \nonumber
\end{align}
Note that the resistance adds in series so $\Rxy = i$ where $i$
is the length of the single path between $x$ and $y$.

The maximum hitting distance is between two vertices whose 
mose recent common ancestor is the root.
For this pair, the hitting time is the same as the commute time
to the root by symmetry.
Consider $a$ and $b$ two leaves whose most recent common ancestor
is the root.
Then
\begin{align}
\E_a [T_b] = \E_a [T_\rho] + \E_\rho [T_a] = 2 (n-1) k \nonumber
\end{align}

and \cref{thm:matthews_up} gives
\begin{align}
\tcov \leq 2(n-1)k\left(1 + \frac{1}{2} + ... + \frac{1}{n} \right). \nonumber
\end{align}

\todo{add lower bound}

\subsection{Spanning Tree Argument}
\cite{AF14}
We introduce the concept of a spanning tree to create
another upper bound for the cover time of a random walk.

\todo{graph example of spanning tree}

A spanning tree $\T$ is a subset of a graph with all edges
in $V(G)$ connected by the minimum number of edges in $E(G)$.
For the connected graphs we consider, the minimum number of edges
needed to connect all vertices is $n-1$.
Formally, $V(\T) = V(G)$ and $E(\T) \subseteq E(G)$
such that $|E(\T)| = n-1$ and $\T$ is connected.

In \cref{sec:commute}, we fixed vertices $a$ and $b$ and called
the effective resistance between them $\Reff$.
We now consider $\Rxy$ the effective resistance between vertices $x$ and $y$.

\begin{theorem}[Spanning Tree Upper Bound] \label{thm:span}
For any spanning tree $\T$ on a weighted graph,
\begin{align}
\tcov \leq C \sum_{x,y \in E(\T)} \Rxy \leq C \sum_{x,y \in E(\T)} 1 / C_{x,y}. \nonumber
\end{align}
In particular for an unweighted graph,
\begin{align}
\tcov \leq 2 |E(G)| (n-1). \nonumber 
\end{align}
\end{theorem}

\begin{proof}
Given a spanning tree $\T$, there exists a vertex $v$ with only one edge.
Consider the path $v_0, v_1, ..., v_{2n-2}$ where $v=v_0=v_{2n-2}$
that traverses each edge in the spanning once in each direction
and covers every vertex.
We construct this path by starting at $v$ and walking along all $n-1$ edges to the
other vertex with only one edge and then coming back to $v$.

The expected time of this particular path on the original graph is certainly less than
or equal to the cover time. Thus
\begin{align}
\tcov &\leq \sum_{i=0}^{2n-3} \E_{v_i}[T_{v_{i+1}}]. \nonumber
\end{align}

We rewrite the sum in terms of commute times,
\begin{align}
\tcov &\leq \sum_{x,y \in E(\T)} \E_x[T_y] + \E_y[T_x]. \nonumber
\end{align}

Using \cref{eqn:commute}, $E_x[T_y] + \E_y[T_x] = C \Rxy$,
\begin{align}
\tcov &\leq C \sum_{x,y \in E(\T)} \Rxy. \nonumber
\end{align}

Observe that if $x \sim y$,
the effective conductance between $x$ and $y$ is at least $C_{x,y}$
since each additional path only increases the flow.
It follows that the effective resistance between $x$ and $y$ is at most $1 / C_{x,y}$ so
\begin{align}
\tcov &\leq C \sum_{x,y \in E(\T)} 1 / C_{x,y}. \nonumber
\end{align}

If the graph is unweighted, then $C_{x,y} = 1$ and $C = 2 |E(G)|$.
Thus
\begin{align}
\tcov \leq 2 |E(G)| (n-1). \nonumber
\end{align}

\end{proof}

@@

\todo{Theorem 2.2.4}

\todo{Example 2.2.5}

\todo{Theorem 3.6.5}

\todo{Example 3.6.6}
